---
title: "ds202_lab4"
author: "Molly Buckler"
date: "10/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(dplyr)

defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
str(biography)
```


##Cleaning the data


##defense cleaning

```{r}
defClean <- defense
defClean[, 1:2] <- sapply(defClean[, 1:2], as.factor)
defClean[, 3:10] <- sapply(defClean[, 3:10], as.numeric)
defClean
```

##Offense cleaning

```{r}
offClean <- offense 
offClean[, 1:2] <- sapply(offClean[, 1:2], as.factor)
offClean[, 3:10] <- sapply(offClean[, 3:10], as.numeric)
offClean
```

##Bio Cleaning
##Hieght is now located in column totalHeight and it is in inches

```{r}
bioClean <- biography 
bioClean[, 1:2] <- sapply(bioClean[, 1:2], as.factor)
newBio <- str_split(bioClean$Height, "-",  simplify = TRUE)
head(newBio)
newBio[, 1:2] <- sapply(newBio[, 1:2], as.numeric)
head(newBio)
bioClean <- merge(bioClean, newBio)
bioClean[, 3:4] <- sapply(bioClean[, 3:4], as.numeric)
bioClean[, 5:7] <- sapply(bioClean[, 5:7], as.factor)
head(bioClean)
bioClean[, 8:9] <- sapply(bioClean[, 8:9], as.numeric)
bioClean <- mutate(bioClean, totalHeight = (12 * V1) + V2)
head(bioClean)

```


##Tidying

```{r}
defClean <- defClean %>%
 pivot_longer(Name, 
              names_to = " Position", 
              values_to = "count")
head(defClean)
defClean %>%
    select_if(is.numeric) %>%
    map_dbl(sum)
```

##The defensive skill that is the least common is Turnover_INT because it only has a count of 6. I found this out because I got the sum of each column which shows how many people have that skill.

```{r}
dat1 <- defClean %>%
  filter(Opponent_Opponent %in% c('Iowa'))
dat1
dat1 %>%
    select_if(is.numeric) %>%
    map_dbl(sum)

dat2 <- defClean %>%
  filter(Opponent_Opponent %in% c('Notre Dame'))
dat2
dat2 %>%
    select_if(is.numeric) %>%
    map_dbl(sum)
```

##ISU had a better defense against notre dame

```{r}
homeTownSplit <- str_split(bioClean$Hometown, ", ",  simplify = TRUE)
head(homeTownSplit)
bioClean <- merge(bioClean, homeTownSplit)
bioClean
bioClean %>% group_by(V2) %>% tally()
```